---
title: "igraph Dataset 1"
output:
  html_document: default
  html_notebook: default
---

Examples pulled from **Network Visualization with R**, [Polnet 2016](http://www.kateto.net/polnet2016) by Katya Ognyanova.

```{r}
suppressPackageStartupMessages({
  library(RColorBrewer)
  library(readr)
  library(dplyr)
  library(igraph)
})
```

```{r}
data_dir <- file.path("inst", "doc", "shiny_logic")
data_dir <- "."
```

For igraph plotting parameters, node options start with `vertex.` and edge options start with `edge.`. Get options with

```{r eval=FALSE}
?igraph.plotting
```

Set node & edge options in two ways:

1. specify them in `plot()` function
2. add them to the `igraph` object

### Basic igraph with Dataset 1

```{r}
nodes <- 
   read_csv(file.path(data_dir,
                      "ShinyLogicNodes.csv"))
links <- 
   read_csv(file.path(data_dir, 
                      "ShinyLogicEdges.csv"))
```

Convert dataset 1 to an igraph object.
The `graph_from_data_frame()` function takes two data frames, `d` (edges of the network) and `vertices` (nodes).
`d` has (at least) two columns with source and target node IDs for each network tie.
`vertices` has (at least) a column of node IDs.
Additional columns in either data frame are interpreted as attributes.

```{r}
(net <- graph_from_data_frame(d=links, vertices=nodes, directed=T)) 
```

Access nodes, edges, and attributes:

```{r}
E(net)
```

```{r}
V(net)
```

Find specific nodes and edges by attribute

```{r}
V(net)[type=="shiny"]
```

Extract an edge list or matrix from igraph networks.

```{r}
tbl_df(as_edgelist(net, names=T))
```

```{r}
as_adjacency_matrix(net, attr="weight")
```

Look at network matrix directly:

```{r}
net[1,]
```

```{r}
net[5,7]
```

Plot graph, removing loops and labels, and reducing arrow size.

```{r}
net <- simplify(net, remove.multiple = F, remove.loops = T) 
```

```{r}
plot(net, 
     edge.arrow.size = 0.4,
     vertex.label = NA)
```

### Annotating a graph

#### Using options to plot

Plot with curved edges (`edge.curved=.1`) and reduce arrow size.
Using curved edges reveals multiple links between two nodes (e.g. links going in either direction, or multiplex links).

```{r}
plot(net, edge.arrow.size=.4, edge.curved=.1)
```

Set node color to `orange` and the border color to hex `#555555`.
Replace the vertex label with the node names stored in `media`.

```{r}
plot(net, edge.arrow.size=.4, edge.curved=0,
     vertex.color="orange", vertex.frame.color="#555555",
     vertex.label=V(net)$media, vertex.label.color="black",
     vertex.label.cex=.7) 
```

Semantic network with only node labels.

```{r}
plot(net, vertex.shape="none", vertex.label=V(net)$media, 
     vertex.label.font=2, vertex.label.color="gray40",
     vertex.label.cex=0.7, edge.color="gray90")
```

#### Using graph attributes

```{r}
# Generate colors based on media type:
colrs <- c("skyblue", "tomato", "slategrey", "gold")
V(net)$color <- colrs[factor(V(net)$logic)]

# Compute node degree (#links) and use it to set node size:
deg <- degree(net, mode="all")
V(net)$size <- deg*3
#V(net)$size <- V(net)$audience.size*0.6

# The labels are currently node IDs.
# Setting them to NA will render no labels:
V(net)$label.color <- "black"
V(net)$label <- NA

# Set edge width based on weight:
#E(net)$width <- E(net)$weight/6

#change arrow size and edge color:
E(net)$arrow.size <- .2
E(net)$edge.color <- "gray80"

plot(net) 
```

Override attributes explicitly in the plot

```{r}
plot(net, edge.color="orange", vertex.color="gray50") 
```

#### Legend

Add legend explaining color meaning.

* `x,y`: legend coordinates
* `pch`: element symbol
* `pt.bg`: background color
* `col`:  border color, 
* `pt.cex`: symbol size
* `bty`: type of box around legend
* `ncol`: number of columns for legend

```{r}
### NOT CHANGED YET.
plot(net) 
legend(x = -1.1, y = -1.1,
       c("Newspaper","Television", "Online News"), 
       pch = 21,
       col = "#777777", pt.bg = colrs, 
       pt.cex = 2.5, 
       bty = "n", ncol = 1)
```

### Highlight Graph Aspects

Color graph edges based on source node color.
Here `ends()` gives us the start and end for each edge in `es`, and 
`names` controls whether `ends()` will return node names or IDs.

```{r}
edge.start <- ends(net, es=E(net), names=F)[,1] # get the "from" node
edge.col <- V(net)$color[edge.start]
plot(net, edge.color = edge.col, edge.curved = .1)
```

Keep only edges with weight higher than network mean using `delete_edges(net, edges)`.

```{r}
cut.off <- mean(links$weight) 
net.sp <- delete_edges(net, E(net)[weight<cut.off])
plot(net.sp) 
```

Plot two tie types (`hyperlink` and `mention`) separately.

```{r}
E(net)$width <- 1.5
net.m <- net - E(net)[E(net)$type=="hyperlink"] # another way to delete edges:
net.h <- net - E(net)[E(net)$type=="mention"]   # using the minus operator
par(mfrow=c(1,2))
plot(net.h, vertex.color="orange", main="Tie: Hyperlink")
plot(net.m, vertex.color="lightsteelblue2", main="Tie: Mention")
```

Force nodes to stay in place for both plots.

```{r}
l <- layout_with_fr(net)
par(mfrow=c(1,2))
plot(net.h, vertex.color="orange", layout=l, main="Tie: Hyperlink")
plot(net.m, vertex.color="lightsteelblue2", layout=l, main="Tie: Mention")
```

Show network communities.

```{r}
par(mfrow=c(1,2))

# Community detection based on label propagation:
clp <- cluster_label_prop(net)
class(clp)

# Community detection returns an object of class "communities" 
# which igraph knows how to plot: 
plot(clp, net)
 
# We can also plot the communities without relying on their built-in plot:
V(net)$community <- clp$membership
colrs <- adjustcolor( c("gray50", "tomato", "gold", "yellowgreen"), alpha=.6)
plot(net, vertex.color=colrs[V(net)$community])
```

#### Highlight Distance to Node(s)

Focus on a particular node or group of nodes. In the example media network, examine spread of information from the `NYT`.

The `distances()` function returns a matrix of shortest paths from nodes listed in the `v` parameter to ones included in the `to` parameter.


```{r}
(dist.from.NYT <- distances(net, v=V(net)[media=="NY Times"], 
                           to=V(net), weights=NA))
```
  
Set colors to plot the distances

```{r}
oranges <- colorRampPalette(c("dark red", "gold"))
col <- oranges(max(dist.from.NYT) + 1)
col <- col[dist.from.NYT + 1]
plot(net, vertex.color = col,
     vertex.label = dist.from.NYT, 
     edge.arrow.size = .6, 
     vertex.label.color = "white")
```

#### Highlight a network path

```{r}
(news.path <- shortest_paths(net, 
                            from = V(net)[media=="MSNBC"], 
                             to  = V(net)[media=="New York Post"],
                             output = "both")) # both path nodes and edges
```

Generate edge color and width and node color variables to plot path.
Awkward, but it works.

```{r}
ecol <- rep("gray80", ecount(net))
ecol[unlist(news.path$epath)] <- "orange"
ew <- rep(2, ecount(net))
ew[unlist(news.path$epath)] <- 4
vcol <- rep("gray40", vcount(net))
vcol[unlist(news.path$vpath)] <- "gold"
```

```{r}
plot(net, 
     vertex.color = vcol, 
     edge.color = ecol, 
     edge.width = ew, 
     edge.arrow.mode = 0)
```

Highlight edges of a vertex, for instance the `WSJ`. For a single node, use `incident()`, for multiple nodes use `incident_edges()`.

```{r}
(inc.edges <- incident(net,  V(net)[media=="Wall Street Journal"], mode="all"))
```

Set colors to plot the selected edges.

```{r}
ecol <- rep("gray80", ecount(net))
ecol[inc.edges] <- "orange"
vcol <- rep("grey40", vcount(net))
vcol[V(net)$media=="Wall Street Journal"] <- "gold"
```

```{r}
plot(net, 
     vertex.color = vcol, 
     edge.color = ecol)
```

Highlight immediate neighbors of a vertex using `neighbors()` to find all nodes one step out from the focal actor. To find the neighbors for multiple nodes, use `adjacent_vertices()`. To find node neighborhoods going more than one step out, use function `ego()` with parameter order set to the number of steps out to go from the focal node(s).

```{r}
(neigh.nodes <- neighbors(net, V(net)[media=="Wall Street Journal"], mode="out"))
```

```{r}
vcol[neigh.nodes] <- "#ff9d00"
plot(net, vertex.color=vcol)
```

Highlight a group of nodes by `mark`ing them (recall network communities).

```{r}
par(mfrow=c(1,2))
plot(net, 
     mark.groups = c(1,4,5,8), 
     mark.col = "#C5E5E7", 
     mark.border = NA)

# Mark multiple groups:
plot(net, 
     mark.groups = list(c(1,4,5,8), c(15:17)),
     mark.col = c("#C5E5E7","#ECD89A"), 
     mark.border = NA)
```

### Other Views of Networks besides Hairballs

Heatmap of network matrix

```{r}
netm <- as_adjacency_matrix(net, attr="weight", sparse=F)
colnames(netm) <- V(net)$media
rownames(netm) <- V(net)$media

palf <- colorRampPalette(c("gold", "dark orange")) 
heatmap(netm[,17:1], Rowv = NA, Colv = NA, col = palf(100), 
        scale="none", margins=c(10,10) )
```

Degree distribution for network.

```{r}
deg.dist <- degree_distribution(net, cumulative=T, mode="all")
plot( x=0:max(deg), y=1-deg.dist, pch=19, cex=1.2, col="orange", 
      xlab="Degree", ylab="Cumulative Frequency")
```
