---
title: "Recla qtl2shiny Demo"
author: "Brian S. Yandell"
date: "2/27/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a demo of the Recla et al. dataset using functionality of [R/qtl2shiny](https://github.com/byandell/qtl2shiny) shown in <http://www.stat.wisc.edu/~yandell/software/qtl2shiny>. We assume that the 
[CCmouse.Rmd](https://github.com/byandell/qtl2shiny/blob/master/inst/doc/CCmouse.Rmd) and 
[Recla.Rmd](https://github.com/byandell/qtl2shiny/blob/master/inst/doc/Recla.Rmd)
scripts have already been run to put data in the `CCmouse` directory.

Here we just attach the `dplyr` and `ggplot2` packages, but explicitly use
elements of the `qtl2`, `qtl2ggplot` and `qtl2pattern` packages. 
The `qtl2feather` package is used implicitly for access to genotype probabilities.

```{r message=FALSE}
library(dplyr)
library(ggplot2)
```

```{r}
project_info <- data.frame(project = "Recla",
                           taxa = "CCmouse",
                           directory = ".",
                           stringsAsFactors = FALSE)
```

```{r}
(project_dir <- file.path(project_info$directory, 
                          project_info$taxa, 
                          project_info$project))
```

```{r}
dir.exists(project_dir)
```

List files for the taxa and for the project.

```{r}
list.files(project_info$taxa)
```

```{r}
list.files(project_dir)
```

```{r}
list.files(file.path(project_dir, "genoprob")) 
```

## Choose Region and Phenotype

```{r}
# See shinyHotspot.R
hotspots <- readRDS(file.path(project_dir, "hotspot.rds"))
```

```{r}
# See shinyHotspot.R
qtl2ggplot::ggplot_scan1(hotspots$scan, hotspots$map)
```

```{r}
# See shinyHotspot.R
(hot_sum <- qtl2pattern::summary_scan1(hotspots$scan, hotspots$map) %>%
  filter(pheno == "all") %>%
  rename(count = lod) %>%
  select(-marker) %>%
  arrange(desc(count)))
```

```{r}
chr_id <- as.character(hot_sum$chr[1])
pos_Mbp <- hot_sum$pos[1]
window_Mbp <- 5
start_val <- pos_Mbp - window_Mbp
end_val <- pos_Mbp + window_Mbp
```

```{r}
# See shinyHotspot.R
autoplot(hotspots$scan, hotspots$map, chr = chr_id)
```

```{r}
# See shinyMain.R
peaks <- readRDS(file.path(project_dir, "peaks.rds")) %>%
  filter(chr == chr_id,
         pos >= start_val,
         pos <= end_val)
```

```{r}
# See shinyPhenos.R
(peaks <- peaks %>%
  select(pheno, chr, pos, lod) %>%
  arrange(desc(lod)))
```

```{r}
pheno_name <- peaks$pheno[1]
```

#### Phenotype Data

```{r}
# See shinyMain.R
(analyses <- readRDS(file.path(project_dir, "analyses.rds")) %>%
  filter(pheno == pheno_name))
```

```{r}
# See shinyMain.R
(pheno_data <- readRDS(file.path(project_dir, "pheno_data.rds")) %>%
   qtl2pattern::pheno_trans(analyses$pheno, 
               analyses$transf,
               analyses$offset,
               analyses$winsorize)) %>% 
  head
```

```{r}
# See shinyMain.R
(covar <- readRDS(file.path(project_dir, "covar.rds")) %>%
  qtl2pattern::get_covar(analyses)) %>%
  head
```

## Genome Scan

```{r}
# See shinyMain.R
kinship <- readRDS(file.path(project_dir, "kinship.rds"))[chr_id]
```

```{r}
# See shinyProbs.R and Recla.Rmd
query_probs <- qtl2pattern::create_probs_query_func(project_dir)
probs_obj <- query_probs(chr_id, start_val, end_val)
```

```{r}
names(probs_obj)
dim(probs_obj$probs)
```

```{r}
# See shinyScanCoef.R
scan_obj <- qtl2pattern::scan1_covar(pheno_data, covar, probs_obj$probs, kinship,
                        analyses)
```

```{r}
autoplot(scan_obj, probs_obj$map)
```

```{r}
summary(scan_obj, probs_obj$map)
```

```{r}
qtl2::find_peaks(scan_obj, probs_obj$map)
```

```{r}
addcovar <- qtl2pattern::covar_df_mx(covar)
eff_obj <- qtl2::scan1coef(probs_obj$probs, pheno_data, kinship, addcovar)
```

```{r}
autoplot(eff_obj, probs_obj$map) +
  geom_vline(xintercept = peaks$pos[1], linetype = "dashed")
```


```{r}
autoplot(eff_obj, probs_obj$map, scan1_output = scan_obj,
         legend.position = "none")
```

## SNP Scans

```{r}
# See shinySNPSetup.R and shinyProbs.R
query_variants <- qtl2::create_variant_query_func(
  file.path(project_info$taxa, "cc_variants.sqlite"))
snpinfo <- query_variants(chr_id, start_val, end_val)
```

```{r}
snpprobs_obj <- qtl2pattern::get_snpprobs(chr_id, pos_Mbp, window_Mbp,
                   pheno_name, 
                   probs_obj$probs,
                   probs_obj$map,
                   snpinfo)
```

```{r}
snp_scan_obj <- qtl2pattern::scan1_covar(pheno_data, covar, snpprobs_obj$snpprobs, kinship, analyses)
```

```{r}
autoplot(snp_scan_obj, snpprobs_obj$snpinfo, show_all_snps = FALSE,
         drop_hilit = 1.5,)
```

```{r}
# See shinySNPSum and shinySNPSetup
top_snps_tbl <- qtl2pattern::top_snps_all(snp_scan_obj,
                        snpprobs_obj$snpinfo,
                        1.5)
summary(top_snps_tbl)
```

#### Genes and Exons

```{r}
# See shinyGeneRegion.R
query_genes <- qtl2::create_gene_query_func(
  file.path(project_info$taxa, "mouse_genes.sqlite"))
rng <- range(top_snps_tbl$pos) + c(-.005, .005)
gene_tbl <- query_genes(chr_id, rng[1], rng[2])
gene_region_tbl <- qtl2pattern::get_genes(chr_id, rng[1], rng[2], gene_tbl)
```

```{r}
autoplot(gene_region_tbl)
```

```{r}
autoplot(gene_region_tbl, top_snps_tbl = top_snps_tbl)
```

```{r}
# See shinyGeneExon.R
gene_exon_tbl <- qtl2pattern::get_gene_exon_snp(top_snps_tbl, gene_tbl)
```

```{r}
(exon_sum <- summary(gene_exon_tbl, top_snps_tbl = top_snps_tbl))
```

```{r}
autoplot(gene_exon_tbl, top_snps_tbl, genes = exon_sum$gene[1])
```

#### SNP Allele Patterns

```{r}
# See shinySNPPattern.R
autoplot(snp_scan_obj, snpprobs_obj$snpinfo, patterns = "hilit", drop_hilit = 1.5)
```

```{r}
# See shinySNPFeature.R
top_feature <- qtl2pattern::merge_feature(top_snps_tbl, snpprobs_obj$snpinfo,
                             snp_scan_obj, 1.5, 0, gene_exon_tbl)
```

```{r}
autoplot(top_feature, pheno_name, "pattern")
```

```{r}
autoplot(top_feature, pheno_name, "consequence")
```

#### Mediation

Going to skip this for now, as it takes another package and is under redesign.

```{r}
peak_mar <- qtl2::find_marker(probs_obj$map, chr_id, pos_Mbp)
```

```{r}
# See shinyMediate.R
all_analyses <- readRDS(file.path(project_dir, "analyses.rds")) %>%
  filter(pheno != pheno_name,
         pheno %in% peaks$pheno)
all_pheno_data <- readRDS(file.path(project_dir, "pheno_data.rds")) %>%
   qtl2pattern::pheno_trans(all_analyses$pheno, 
               all_analyses$transf,
               all_analyses$offset,
               all_analyses$winsorize)
annot <- dplyr::left_join(
  peaks %>% filter(pheno %in% all_analyses$pheno),
  all_analyses,
  by = "pheno")
annot$driver <- qtl2::find_marker(probs_obj$map, chr_id, annot$pos)
med_ls <- 
  list(comediators = qtl2pattern::pheno_trans(all_pheno_data,
                                              annot$pheno,
                                              annot$transf,
                                              annot$offset,
                                              annot$winsorize),
       annot = dplyr::rename(annot, 
                             id = pheno,
                             biotype = pheno_type),
       cov_med = addcovar,
       phe_type = (dplyr::filter(peaks, pheno == pheno_name))$pheno_type[1])
```

```{r}
geno_max <- subset(probs_obj$probs, chr = chr_id, mar = peak_mar)[[1]][,,1]
```

```{r}
driver_med <- probs_obj$probs[[chr_id]]
```

```{r}
mediate_obj <- CausalMST::mediate1_test(med_ls, geno_max, pheno_data,
                           kinship[[1]], addcovar, med_ls$cov_med,
                           driver_med,
                           test = "wilc", pos = pos_Mbp)
```

```{r}
autoplot(mediate_obj)
```

```{r}
# See shinyScatter.R

```

## Gene Action

```{r}
# See shinyDiplo.R and shinyProbs.R
pairprobs_obj <- query_probs(chr_id, start_val, end_val, allele = FALSE)
```

```{r}
snppairprobs_obj <- qtl2pattern::get_snpprobs(chr_id, pos_Mbp, window_Mbp,
                   pheno_name, 
                   pairprobs_obj$probs,
                   pairprobs_obj$map,
                   snpinfo)
```

```{r}
snppair_scan_obj <- qtl2pattern::scan1_covar(pheno_data, covar, 
                                snppairprobs_obj$snpprobs, kinship, analyses)
```

```{r}
autoplot(snppair_scan_obj, snppairprobs_obj$snpinfo, show_all_snps = FALSE,
         drop_hilit = 1.5,)
```

```{r}
# See shinySNPPattern.R
autoplot(snppair_scan_obj, snppairprobs_obj$snpinfo, patterns = "hilit", drop_hilit = 1.5)
```

```{r}
top_snppairs_tbl <- 
  qtl2pattern::top_snps_all(snppair_scan_obj,
               snppairprobs_obj$snpinfo,
               1.5)
```

```{r}
(patterns <- summary(top_snppairs_tbl))
```

```{r}
# See shinyPattern.R
scan_pat <- qtl2pattern::scan_pattern(pairprobs_obj$probs,
                            pheno_data,
                            kinship, addcovar,
                            pairprobs_obj$map,
                            patterns)
```

```{r}
pattern_choices <- qtl2pattern::sdp_to_pattern(patterns$sdp, LETTERS[1:8])
```

```{r}
autoplot(scan_pat, pairprobs_obj$map, "lod", pattern_choices)
```

## Multiple Phenotypes

```{r}
pheno_names <- peaks$pheno[1:2]
```

Assume for now covariates are the same. The routine `qtl2pattern::scan1_covar` helps sort out covariates that change among phenotypes.

```{r}
# See shinyMain.R
(analyses <- readRDS(file.path(project_dir, "analyses.rds")) %>%
  filter(pheno %in% pheno_names))
```

```{r}
# See shinyMain.R
(pheno_data <- readRDS(file.path(project_dir, "pheno_data.rds")) %>%
   qtl2pattern::pheno_trans(analyses$pheno, 
               analyses$transf,
               analyses$offset,
               analyses$winsorize)) %>% 
  head
```
```{r}
# See shinyScanCoef.R
scan_obj <- qtl2pattern::scan1_covar(pheno_data, covar, probs_obj$probs, kinship,
                        analyses)
```

```{r}
autoplot(scan_obj, probs_obj$map, 1:2)
```

```{r}
snp_scan_obj <- qtl2pattern::scan1_covar(pheno_data, covar, snpprobs_obj$snpprobs, kinship, analyses)
```

```{r}
autoplot(snp_scan_obj, snpprobs_obj$snpinfo, 1:2, show_all_snps = FALSE,
         drop_hilit = 1.5, facet = "pheno")
```

```{r}
# See shinySNPPattern.R
autoplot(snp_scan_obj, snpprobs_obj$snpinfo, 1:2, 
         patterns = "hilit", drop_hilit = 1.5, facet = "pheno")
```
