---
title: "visNetwork Dataset 1"
output:
  html_document: default
  html_notebook: default
---

Examples pulled from **Network Visualization with R**, [Polnet 2016](http://www.kateto.net/polnet2016) by Katya Ognyanova.

```{r}
suppressPackageStartupMessages({
  library(RColorBrewer)
  library(readr)
  library(dplyr)
  library('visNetwork') 
})
```

Here use [visNetwork](http://datastorm-open.github.io/visNetwork/) package.
visNetwork enables setting graphic properties as node or edge attributes by adding them as data columns before calling `visNetwork()`. See available options with:

```{r eval=FALSE}
?visNodes
?visEdges
```

For more information, see:

```{r eval=FALSE}
?visOptions # available options 
?visLayout  # available layouts
?visGroups  # using node groups
?visLegend  # adding a legend
```

### Basic visNetwork with Dataset 1

```{r}
nodes <- 
   read_csv(system.file("extdata", "ShinyLogicNodes.csv", package = "qtl2shiny")) %>%
  mutate(id = node) %>%
  select(-node)
links <- 
   read_csv(system.file("extdata", "ShinyLogicEdges.csv", package = "qtl2shiny")) %>%
  mutate(from = source,
         to = target) %>%
  select(-source,-target)
```

Open dataset 1 with visNetwork.

```{r}
visNetwork(nodes, links, width="100%", height="400px", main="Network!")
```

Changing some node parameters:

* shape (ellipse, circle, database, box, text, image, circularImage, diamond, dot, star, triangle, triangleDown, square, icon)
* color.background (node color)
* color.border (frame color)
* color.highlight (color on mouse click)
* color.hover (color on mouseover)

```{r}
nodes$shape <- "dot"  
nodes$shadow <- TRUE # Nodes will drop shadow
nodes$title <- nodes$id # Text on click
nodes$label <- nodes$logic # Node label
#nodes$size <- nodes$audience.size # Node size
nodes$borderWidth <- 2 # Node border width

nodes$color.background <- c("skyblue", "tomato", "slategrey", "gold")[factor(nodes$logic)]
nodes$color.border <- "black"
nodes$color.highlight.background <- "orange"
nodes$color.highlight.border <- "darkred"

visNetwork(nodes, links)
```

Set up hover option.

```{r}
nodes$color.hover.background <- "green"
visNetwork(nodes, links) %>%
 visInteraction(hover = TRUE)
```

Change visual properties of edges.

```{r}
#links$width <- 1 + links$weight/8 # line width
links$color <- "gray"    # line color  
links$arrows <- "middle" # arrows: 'from', 'to', or 'middle'
links$smooth <- FALSE    # should the edges be curved?
links$shadow <- FALSE    # edge shadow

visNetwork(nodes, links)
```

`visNetwork` offers a number of other options in the `visOptions()` function. For instance, highlight all neighbors of a selected node (`highlightNearest`), or add a drop-down menu to select groups of nodes (`selectedBy`). The groups are based on data columns such as `type label`.

```{r}
## nodes must be a data frame, not tbl_df.
visNetwork(as.data.frame(nodes), links) %>%
  visOptions(highlightNearest = TRUE, 
             selectedBy = "type.label")
```

